# Sculpting data for fun and profit
## Libraries
```{r}
    library(tidyverse)
    library(sf)
    library(ggplot2)
  library(tmap)
```

## Import the data
```{r}
    weca = st_as_sf(st_read("./data/weca.gpkg"))
```

## 1: Looking at the data

### 1.1: Filton temperatures
```{r}
    library(readr)
    temps = read_csv("./data/bristol_yearly_temps.csv")
    temps
```
This data is bad because variables (max/min/median) are in one singular column, so it doesn't follow the rules. This is easy to fix.

### 1.2: Fixing Filton's temperatures
It's as easy as a single pivot
```{r}
    tidyTemps = pivot_wider(temps, 
        c("year","stat"),
        names_from = "stat",
        values_from = "degrees_celsius"
    )
    head(tidyTemps)
```
All three rules are now met, seporate columns for different datatypes, a single row per record and a single datapoint per cell

### 1.3: WECA has a problem
As per usual, WECA just isn't up to scratch. In this case, it's got datapoints in the headers. That just won't do. We must abolish WECA('s datapoints).

## 2: Tidying the data
### 2.1: Tidying up a WECA
```{r}
    tidyWeca = pivot_longer(weca,5:97,
    names_to = c("month", "year"), values_to = "housePrice", names_prefix = "price_", names_sep = "_")
tidyWeca$year = as.numeric(tidyWeca$year)
```
In all cases, the date (month/year) is in its own column now, yay

### 2.2: A boxplot
```{r}
ggplot(tidyWeca, aes(y=housePrice, x=year, color=la_name))+geom_boxplot()
```
### 2.3: More summary stats

#### Average price per year
```{r}
appy = group_by(tidyWeca, year) %>% summarise(mean=mean(housePrice, na.rm=T))
```
#### Average price per LA per year
```{r}
applay = group_by(tidyWeca, year ,la_name) %>% summarise(mean=mean(housePrice, na.rm=T))
```

#### Median price per LSOA per year since 2008
```{r}
mplsoa2008 = filter(tidyWeca, year >=2008) %>% group_by(lsoa_name) %>% summarise(mean=median(housePrice, na.rm=T))
```

#### 2.c: Percentage change map
```{r}
# get the average price by year for each lsoa
appylsoa = filter(tidyWeca, year == 2008 | year==2018) %>% group_by(year, lsoa_name) %>% summarise(mean=mean(housePrice, na.rm=T))
```

```{r}
pcm = group_by(appylsoa, lsoa_name) %>% summarise(pc = ((last(mean)+first(mean))/first(mean)))
```

```{r}
qtm(pcm, fill=pc)
```