# 4: things
## Tufty is back
I do't like tufty's style, but it is pretty good

## Principle of proportional ink
- The things you spend ink on are the things people will look at
- Anything that goes on the plot must be useful
- Stuff that has more ink is more important
- Don't include unneded lines
- colour is good, but must be clear
- chrome is as minimal as possible

### redundant coding
- one variable shows as two or more visuals
- makes the point even clearer
- we can perhaps even use colour in facets, which i don't think makes sense but we're doing it anyway

### colour (theory)
#### ramps
- a set of colours

##### sequential
- from one colour to another
- two point gradient
- low to high intensity according to colour theory

##### diverging
- three point gradient
- high to low to high intensity

##### qualitative/categorical
-

cartocolours looks very pretty



# The actual exercise
```{r}
library(tidyverse)
library(ggrepel)
library(hexbin)
```
```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(label = model), data = best_in_class)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_point(size = 3, shape = 1, data = best_in_class) +
  ggrepel::geom_label_repel(aes(label = model), data = best_in_class)
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15, 40, by = 5))

presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_x_date(NULL, breaks = presidential$start, date_labels = "'%y")

    base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))

base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right") # the default

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1, override.aes = list(size = 4)))
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x'?

ggplot(diamonds, aes(carat, price)) +
  geom_bin2d()

ggplot(diamonds, aes(log10(carat), log10(price))) +
  geom_bin2d()
```


```{r}
    ggplot(mpg, aes(hwy, cty)) +
    geom_point(aes(colour=class)) +
    geom_smooth(method="lm", data=filter(mpg, class!="2seater"), se=F) +
    labs(
        title="Fuel Economy under different conditions",
        x="City Driving / mpg", 
        y="Highway Driving / mpg"
    )
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_colour_discrete() +
  scale_colour_brewer(palette = "Set1")
```

```{r}
    presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```

```{r}
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)
ggplot(df, aes(x, y)) +
  geom_hex() +
  coord_fixed()

ggplot(df, aes(x, y)) +
  geom_hex() +
  viridis::scale_fill_viridis() +
  coord_fixed()

```

```{r}
ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_fill_gradient(low = "white", high = "red") +
  coord_fixed()
```

```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_colour_manual(values = c(Republican = "red", Democratic = "blue")) +
    scale_y_continuous()
```